Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over convex domains and domains with shellable triangulations

### 2024-10-04 draft.tex
shape measures

>>>>

We shift gears and move from a combinatorial study of triangulations to an analytical study:
we introduce quantitative measures for the geometric quality of triangulations. 
These indicate of the regularity of simplex shapes. 
All these quantities have in common that they can be computed from purely local information. 
Most of such \emph{shape measures} in the literature are based on geometric features, 
but it will also be helpful to use indicators based on affine transformations associated with the mesh. %;this is because the properties of these transformations are what shows up in our estimates. 
While there are many equivalent shape measures, we use the ones that most immediate appear in our proofs.

### 2024-10-04 main.tex
shape measures

@@

We shift gears and move from a combinatorial study of triangulations to an analytical study:
we introduce quantitative measures for the geometric quality of triangulations. 
These indicate of the regularity of simplex shapes. 
All these quantities have in common that they can be computed from purely local information. 
Most of such \emph{shape measures} in the literature are based on geometric features, 
but it will also be helpful to use indicators based on affine transformations associated with the mesh. %;this is because the properties of these transformations are what shows up in our estimates. 
While there are many equivalent shape measures, we use the ones that most immediate appear in our proofs.

\mwl{MWL: The different geometric quantities in this section should really be just what is needed later. Any shape measure should be locally measurable and close to what appears in the right-hand sides of the main results. Any inequalities here should be as tight as possible. }

We write $\diam(T)$ and $\vol(T)$ for the diameter and $n$-dimensional volume of any $n$-simplex $T$.
Moreover, $\height(T)$ refers to the smallest height of any of the vertices of the simplex $T$,
where the height of a vertex is defined as the distance to the affine span of its opposing face.

For the purpose of the usual scaling arguments, the \emph{$n$-dimensional reference simplex} $\Delta^n \subseteq \bbR^n$ is the convex closure of the origin and the $n$ canonical unit vectors.

Given a matrix $A \in \bbR^{n \times n}$, we let $\sigma_{1}(A) \geq \dots \geq \sigma_{n}(A) \geq 0$ denote the descending sequence of its singular values, 
where $\sigma_{\max}(A) := \sigma_{1}(A)$ is the largest and $\sigma_{\min}(A) := \sigma_{n}(A)$ is the smallest. 
We recall that the $\ell^{2}$ operator norm of $A$, written $\| A \|_{2}$, agrees with $\sigma_{\max}(A)$,
and if $A$ is invertible, then $\|A^{-1}\| = \sigma_{\min}(A)^{-1}$.
% Another norm on matrices, less common but useful for us, is the \emph{maximum column norm} $\|A\|_{\maxcol}$, which is defined as the largest $\ell^{2}$-norm of any column of $A$.

Whenever $T$ is any $n$-dimensional simplex $T$,
we define the \textit{geometric shape measure} $\geometricshapemeasure(T)$,
the \textit{aspect shape measure} $\aspectratio(T)$,
and 
the \textit{algebraic shape measure} $\algebraicshapemeasure(T)$
by 
\begin{align}\label{math:shapemeasure}
    \geometricshapemeasure(T)
    := 
    \frac{ \diam(T)^{n} }{ n! \vol(T) }
    ,
    \qquad 
    \aspectratio(T)
    := 
    \frac{ \diam(T) }{ \height(T) }
    ,
    \qquad 
    \algebraicshapemeasure(T)
    := 
    \sup_{ \varphi : \Delta^n \rightarrow T } 
    \frac{ \sigma_{\max}( \Jacobian \varphi ) }{ \sigma_{\min}( \Jacobian \varphi ) }
%     \sup_{ \varphi : \Delta^n \rightarrow T } \matnorm{\Jacobian \varphi}_{\maxcol} \matnorm{ \Jacobian \varphi^{-1} }_{\maxcol}
    ,
\end{align}
where the last supremum is taken over all affine transformation from the reference $n$-simplex onto the $n$-simplex $T$. 
% DONE: Although this is fairly clear upon reading, we could specify in word that $\geometricshapemeasure(T)$ is (equivalent to) the usual shape regularity parameter, whereas $\volumeratio(\mathcal T)$ is the maximal ratio of volumes between neighboring elements.

When $\calT$ is an $n$-dimensional simplicial complex, we naturally introduce 
\begin{align}\label{math:shapemeasure:triangulation}
    \geometricshapemeasure(\calT) := \sup_{ T \in \subsimplex_{n}(\calT) } \geometricshapemeasure(T)
    ,
    \quad 
    \aspectratio(\calT) := \sup_{ T \in \subsimplex_{n}(\calT) } \aspectratio(T)
    ,
    \quad 
    \algebraicshapemeasure(\calT) := \sup_{ T \in \subsimplex_{n}(\calT) } \algebraicshapemeasure(T)
    .
\end{align}
We call these the geometric, aspect, and algebraic shape measure, respectively, of the triangulation.

We now gather a few relationships between geometric and algebraic entities and compare the different shape measures of a single simplex.

\begin{lemma}\label{lemma:measurerelationships}
    Let $T$ be an $n$-simplex and let $\varphi : \Delta^{n} \rightarrow T$ be an affine diffeomorphism from the reference $n$-simplex. Then 
    \begin{gather*}
        \matnorm{\Jacobian \varphi}
        \leq 
        \Ceins{n} \cdot \diam(T) % DONE: can this improved?
        ,
        \qquad 
        \matnorm{ \Jacobian \varphi^{-1} }
        \leq 
        \Czwei{n} \cdot 
        \geometricshapemeasure(T) 
        \diam(T)^{-1}
        ,
        \\
        \algebraicshapemeasure(T)
        \leq 
        \Cdrei{n}
        % \sqrt{ \frac{n}{ (n-1)^{n-1} } }
        \geometricshapemeasure(T)
        ,
        \qquad 
        %         \geometricshapemeasure(T)^{\frac 1 n}
        %         \leq 
        \aspectratio(T)
        \leq 
        \geometricshapemeasure(T)
        .
    \end{gather*}
    Here,
    \begin{align*}
        \Ceins{n} \leq \sqrt{n},
        \quad 
        \Czwei{n} \leq          \left( \frac{n-1}{n} \right)^{\frac{2}{n-1}}, 
        \quad 
        \Cdrei{n} \leq \Ceins{n}\Czwei{n} %\sqrt{n} \left( \frac{n}{n-1} \right)^{\frac{n-1}{2}}
        . 
    \end{align*}
\end{lemma}
\begin{proof}
%     The columns of the matrix $\Jacobian \varphi^{-1}$ are the gradients of the barycentric coordinates of the vertices of $T$,
%     except for the vertex $\varphi(0)$. It immediately follows that 
%     \begin{align*}
%         \matnorm{ \Jacobian \varphi^{-1} }_{2} \leq \sqrt{n} \cdot \height(T)^{-1}
%         .
%     \end{align*}
%     The smallest height in the reference simplex $\Delta^{n}$ is $1/\sqrt{n}$. 
%     
    We easily verify the first inequality. To prove the second inequality, let $\sigma_1 \geq \dots \geq \sigma_n \geq 0$ denote the singular values of the Jacobian $\Jacobian\varphi$ in descending order.
    The Jacobian determinant is the product of these singular values. 
    By a result of Hong and Pan \cite{hong1992lower},
    \begin{align*}
        \sigma_{n}
        &\geq 
        \left( \frac{n-1}{n} \right)^{\frac{n-1}{2}}
        \frac{ \det(\Jacobian\varphi) }{ \diam(T)^{n-1} }
        = 
        \left( \frac{n-1}{n} \right)^{\frac{n-1}{2}}
        \frac{ n! \vol(T) }{ \diam(T)^{n-1} }
        .
    \end{align*}
%     To get the second inequality, we multiple the last term by $\sigma_{1} / ( \sqrt{n} \diam(T) ) \leq 1$.
    The second inequality follows by taking the reciprocal. 
    The definition of the condition number now gives the third inequalities. 
    % To get the third inequality, we divide both sides by $\sigma_{1}$ and use $\sigma_{1} \leq \sqrt{n} \diam(T)$.
    Finally, 
    if $F$ is the face of $T$ opposite to the vertex with the smallest height, then 
    \begin{align*}
        \frac{ \diam(T) }{ \height(T) }
        \leq 
        \frac{ \diam(T) \vol(F) }{ n \vol(T) }
        \leq 
        \frac{ \diam(T)^{n} }{ n! \vol(T) }
        .
    \end{align*}
%     Since $n! \vol(T)$ is bounded from below by the product of any $n$ different heights of $T$, it follows that 
%     \begin{align*}
%         \frac{ \diam(T)^{n} }{ n! \vol(T) }
%         \leq 
%         \frac{ \diam(T)^{n} }{ \height(T)^{n} }
%         .
%     \end{align*}
    That shows the fourth and last of the inequalities. 
\end{proof}

\begin{remark}
    Whenever $T$ is an $n$-dimensional simplex, the ratio $\geometricshapemeasure(T)$ measures the ``shape quality'' of a simplex and is instance of a so-called \emph{shape measure}.
    In our convention, the reference simplex has unit shape measure.
    For example, $\geometricshapemeasure(T)=1$ whenever $T$ is a line segment, an isosceles right-angled triangle, or the generalizations of these reference simplices to higher dimensions. $\geometricshapemeasure(T)$ is slightly bigger than one for equilateral simplices and blows up with distortion.

% Whenever $T$ is an $n$-dimensional simplex, the ratio $\diam(T)^{n} / \vol(T)$ measures the ``shape quality'' of a simplex and is instance of a so-called \emph{shape measure}.
    % We can also rescale that by a factor of $1/n!$, as suggested from the above. 
    The geometric ratio $\geometricshapemeasure(T)$ equals what is known as \emph{fatness} in differential geometry~\cite{cheeger1984curvature} and is precisely the reciprocal of the \emph{fullness} discussed by Whitney~\cite{whitney2012geometric}. 
    The \emph{thickness} of a simplex is the ratio of its smallest height to its diameter~\cite{munkres2016elementary}. \mwl{Update this entire paragraph.}
    Alternative shape measures have been used throughout the literature of numerical analysis and computational geometry to quantify the quality of simplices.
    One example, commonly referenced in finite element literature, is the ratio of the diameter and the radius of the largest inscribed circle of a simplex 
% TODO     (see~\cite[p.61, Definition 5.1]{braess2001finite}, ~\cite[p.97, Definition (4.2.16)]{brenner2008mathematical},~\cite[Definition~11.2]{ern2021finite}). 
\end{remark}

We are interested in the maximal ratio of volumes between \mwl{face?} neighboring $n$-simplices, written $\volumeratio(\calT)$,
and the ratio of intersecting simplices, written $\diameterratio(\calT)$. Formally, 
\begin{align}\label{math:volumeratio}
    \volumeratio(\calT) 
    := 
    \sup_{ \substack{ T, T' \in \subsimplex_{n}(\calT) \\ T \cap T' \in \subsimplex_{n-1}(\calT) } } 
    \frac{ \vol(T) }{ \vol(T') }
    ,
\end{align}
\begin{align}\label{math:diameterratio}
    \diameterratio(\calT) 
    := 
    \sup_{ \substack{ T, T' \in \subsimplex_{n}(\calT) \\ T \cap T' \neq \emptyset } }
    \frac{ \diam(T) }{ \diam(T') }
    .
\end{align}
Finally, whenever $T, T'$ are two $n$-simplices that share a common face $F$ of codimension $1$, we let $\Xi_{T,T'} : T \rightarrow T'$ denote the affine diffeomorphism 
that preserves $F$. We then define 
\begin{align}\label{math:reflectionmeasure}
    \reflectionestimate(\calT) 
    := 
    \sup_{ \substack{ T, T' \in \subsimplex_{n}(\calT) \\ T \cap T' \in \subsimplex_{n-1}(\calT) } }
    \matnorm{ \Jacobian \Xi_{T,T'} }
\end{align}
to be the maximum of the operator norm of the Jacobian of any such diffeomorphism. 
This indicator quantifies how much reflection across the shared face distorts the geometry.

% and $\varphi : \Delta^{n} \rightarrow T$ is an affine diffeomorphism from the reference simplex,
% we define the quantity 
% \begin{align*}
%     \geometricshapemeasure(T)
%     := 
%     \frac{ \diam(T)^{n} }{ n! \vol(T) }
%     .
% \end{align*}

\begin{lemma}\label{lemma:volumecomparison}
    Let $T_1$ and $T_2$ be two $n$-simplices that share a common face $F$. Then 
    \begin{align*}
        \diam(T_1)
        \leq 
        % \frac{ \diam(T_1)^{n} }{ n! \vol(T_1) }
        % \diam(F)
        % = 
        \geometricshapemeasure(T_1)
        \diam(F)
        ,
        \qquad 
        \frac{ \vol(T_1) }{ \vol(T_2) }
        \leq 
        % \frac{\diam(T_1)^{n}}{n! \cdot \vol(T_1)} 
        % \cdot 
        % \frac{\diam(T_2)^{n}}{n! \vol(T_2)} 
        % = 
        \aspectratio(T_1) \aspectratio(T_2)
        .
    \end{align*}
    If $\Xi : T_1 \rightarrow T_2$ is an affine diffeomorphism that is the identity over $F$, then 
    \begin{align*}
        \matnorm{ \Jacobian \Xi }
        \leq 
        ( 1 + \aspectratio(T_2) ) \aspectratio(T_1)
        % \Cvier{n} \geometricshapemeasure(T_2) \geometricshapemeasure(T_1) 
        .
    \end{align*}
%     Here, $\Cvier{n} \leq \Ceins{n} \Czwei{n}$. 
\end{lemma}

% Together with analogous estimates that swap the roles of $h_1$ and $h_2$, we find 
% \begin{align*}
%     \max\left( \frac{h_1}{h_2}, \frac{h_2}{h_1} \right)
%     \leq 
%     \frac{\diam(T_1)^{n}}{n! \cdot \vol(T_1)} 
%     \cdot 
%     \frac{\diam(T_2)^{n}}{n! \vol(T_2)} 
%     .
% \end{align*}
% DONE: REMOVE THE FACTORIAL?
% DONE: I am not 100% sure what that means. Is really the heights of F? I imagine this is rather the heights of simplices T_1 and T_2, isn't? Maybe the wording you use is standard and I simply don't know about it, if so, please ignore the comment! (I mean, I am probably being too pedantic here, what you mean is completely clear from the next sentence).
% DONE: clarify the first of the following inequalities

%     \begin{align*}
%         \diam(F) \geq h_S = n \frac{ \vol(T_1) }{ \vol(S) } \geq n \frac{ \vol(T_1) }{ \diam(T_1)^{n} } \diam(T_1).
%     \end{align*}
%     Thus we establish the first estimate:

%     Let $h$ be the height of some face $F'$ other than $F$ in $T_1$. Then $h \leq \diam(F)$, and with similar arguments as before: 
%     \begin{align*}
%         \diam(F)
%         \geq 
%         h
%         =
%         n
%         \frac{ \vol(T_1) }{ \vol(F') }
%         \geq 
%         \frac{ n \vol(T_1) }{ \diam(T_1)^{n} }
%         \cdot 
%         \diam(T_1)
%         .
%     \end{align*}
%     That shows the second estimate.
    The diameter of $F$ is at least as large as the height $h_S$ of some other facet $S$ of $T_{1}$. 
%     Recursive application of the volume formula shows $n! \cdot \vol(T) = h_S h_2 \cdots h_n$,
%         n! \cdot \vol(T) = \frac{1}{n!} h_S h_2 \cdots h_n,
%     where $h_S \leq \diam(F)$ and $h_2, \cdots, h_n \leq \diam(T_{1})$. 
    Now, 
%        \diam(T) \frac{ n! \vol(T) }{ \diam(T)^{n} } \leq h_S \leq \diam(F).
        \diam(T_{1}) \aspectratio(T_{1})^{-1} \leq \height(T_{1}) \leq h_S \leq \diam(F).
    The first estimate follows. 
    As for the second estimate,
    let $h_1$ and $h_2$ be the heights of $F$ in the simplices $T_1$ and $T_2$, respectively. 
    By the volume formula for simplices, $\vol(T_1) = h_1 \vol(F)/n$ and $\vol(T_2) = h_2 \vol(F)/n$, and thus $\vol(T_1)/\vol(T_2) = h_1/h_2$. Thus follows the second estimate: 
        \frac{ h_1 }{ h_2 }
        \frac{ \diam(T_1) }{ h_2 }
        \aspectratio(T_{1}) \frac{ \diam(F) }{ h_2 }
        \aspectratio(T_{1}) \aspectratio(T_{2})
    On the other hand, 
        % h_1 &\leq \diam(T_1) \leq \frac{\diam(T_1)^{n}}{n \cdot \vol(T_1)} \diam(F),
%         h_2 &= n \frac{\vol(T_2)}{\vol(F)} \geq n \cdot \frac{\vol(T_2)}{\diam(T_2)^{n}} \diam(T_2) \geq n \cdot \frac{\vol(T_2)}{\diam(T_2)^{n}} \diam(F)
        h_1 \leq \aspectratio(T_{1}) \diam(F),
        h_2 \geq \aspectratio(T_{2})^{-1} \diam(T_2)\geq \aspectratio(T_{2})^{-1} \diam(F)
    The second estimate follows. 
    Lastly, $\Xi : T_1 \rightarrow T_2$ be as above
    can be written as the composition $\Xi = \phi_{2} \phi^{-1}_{1}$ for some affine transformations $\phi_1 : \Delta^{n} \rightarrow T_1$ and $\phi_2 : \Delta^{n} \rightarrow T_2$. 
    Lemma~\ref{lemma:measurerelationships} and the first estimate imply 
        \matnorm{ \Jacobian \Xi }
        \matnorm{ \Jacobian\phi_{2} } \cdot \matnorm{ \Jacobian\phi_{1}^{-1} }
        \Ceins{n}^{2} \diam(T_2) \height(T_1)^{-1}
        \Ceins{n}^{2} \aspectratio(T_2) \diam(F) \height(T_1)^{-1}
        \Ceins{n}^{2} \aspectratio(T_2) \aspectratio(T_1)
%         \matnorm{ \Jacobian \Xi }
%         &
%         \Ceins{n}
%         \Czwei{n}
%         \geometricshapemeasure(T_2)
%         \diam(T_1)
%         \diam(T_2)^{-1}
%         \Ceins{n}
%         \Czwei{n}
%         \geometricshapemeasure(T_2)
%         \geometricshapemeasure(T_1) 
%         \diam(F)
%         \diam(T_2)^{-1}
%         \Ceins{n}
%         \Czwei{n}
%         \geometricshapemeasure(T_2)
%         \geometricshapemeasure(T_1) 
    All the desired results thus follow.
    Lastly, let $\Xi : T_1 \rightarrow T_2$ be as stated. 
    Writing $h_0$ for the unit normal in direction $z_1$, we have 
        \Xi(x) 
        - \frac{ \langle h_0, x \rangle }{ \langle h_0, z_1 \rangle } z_1
        + \frac{ \langle h_0, x \rangle }{ \langle h_0, z_1 \rangle } z_2
        x \in \bbR^{n}
    Its Jacobian is the rank-one perturbation of the linear projection along $z_1$ onto the hyperplane spanned by $F$. 
    By the Xu-Zikanatov estimate,
        \matnorm{ \Jacobian \Xi }
        \frac{ \diam(T_1) }{ \height(T_1) }
        \frac{ \diam(T_2) }{ \height(T_1) }
        \left( 1 + \aspectratio(T_2) \right)
        \aspectratio(T_1)
    The desired estimates are shown.
    \mwl{Tighter estimate via geometry}
    Matrix inverse lemma:
        A ( A + (z_2-z_1) e_1^t )^{-1}
        \left( 1 + e_1^{t} A^{-1} (z_2-z_1) \right)^{-1}
        (z_2-z_1) e_1^t \cdot A^{-1}
        \left( e_1^{t} A^{-1} z_2 \right)^{-1}
        (z_2-z_1) e_1^t \cdot A^{-1}
    Putting in a multiple of $\alpha z_1$, we find 
        \alpha z_1
        \left( e_1^{t} A^{-1} z_2 \right)^{-1}
        (z_2-z_1) 

@@

\begin{remark}
    Lemma~\ref{lemma:volumecomparison} fully controls the shape quantity $\volumeratio(\calT)$ \cye{of~\eqref{math:volumeratio}} but not $\diameterratio(\calT)$ \cye{of~\eqref{math:diameterratio}}. 
    Lemma~\ref{lemma:volumecomparison} fully controls the shape quantity $\volumeratio(\calT)$ but not $\diameterratio(\calT)$. 
    As seen above, we can bound the ratio of diameters and volumes of simplices that share at least one face in terms of the shape measure. 
    If we have an upper estimate for the number of simplices sharing a given vertex, 
    we can also bound the ratios of diameters and volumes of simplices sharing only a single common vertex. 
\end{remark}
%%%%%%%% END HERE

@@

\begin{corollary}
    Under the assumptions of Theorem~\ref{theorem:poincarefriedrichsestimate:grad},
    whenever $T_0, T_1, \dots, T_m$ is a face-path of $n$-simplices in $\calT$,
    we have 
    \begin{align*}
        \| w \|_{L^{p}(T_m)}
        &\leq 
        \sum_{\ell=1}^{m} 
        \volumeratio^{\frac{m-\ell}{p}}
        \min\left( 
        \left( 1 + \volumeratio(\calT)^{\frac q p} \right)^{\frac 1 q}
        C_{{\PF},\Minako_{F_\ell},p} 
        ,
        \left( 1 + \volumeratio(\calT)^{\frac q p} \xi(\calT)^{q} \right)^{\frac 1 q}
        C_{\PF,T_{\ell},F_{\ell},p} 
        \right)
        \| \nabla u \|_{L^{p}(\Minako_{F_\ell})}
        \\&
        \qquad 
        \qquad 
        +
        \volumeratio^{\frac m p}
        C_{{\PF},T_{0},p}
        \| \nabla u \|_{L^{p}(T_{0})}
        .
    \end{align*}
\end{corollary}
\begin{proof}
    This follows by repeated application of the recursive estimate in Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.
\end{proof}

>>>>

\paragraph{Notation}

\mwl{Fix this}
Given a matrix $A \in \bbR^{n \times n}$, we let $\sigma_{1}(A) \geq \dots \geq \sigma_{n}(A) \geq 0$ denote the descending sequence of its singular values, 
where $\sigma_{\max}(A) := \sigma_{1}(A)$ is the largest and $\sigma_{\min}(A) := \sigma_{n}(A)$ is the smallest. 
We recall that the $\ell^{2}$ operator norm of $A$, written $\| A \|_{2}$, agrees with $\sigma_{\max}(A)$,
and if $A$ is invertible, then $\|A^{-1}\| = \sigma_{\min}(A)^{-1}$.
% Another norm on matrices, less common but useful for us, is the \emph{maximum column norm} $\|A\|_{\maxcol}$, which is defined as the largest $\ell^{2}$-norm of any column of $A$.

@@

We introduce several quantities that measure the regularity of a triangulation. 
% We quantify the quality of a triangulation via so-called \emph{shape measures}. 
These have in common that they can be computed from purely local information.

We write $\diam(T)$ and $\vol(T)$ for the diameter and $n$-dimensional volume of any $n$-simplex $T$.
Moreover, $\height(T)$ refers to the smallest height of any of the vertices of the simplex $T$,
where the height of a vertex is defined as the distance to the affine span of its opposing face.
For the purpose of the usual scaling arguments, the \emph{$n$-dimensional reference simplex} $\Delta^n \subseteq \bbR^n$ is the convex closure of the origin and the $n$ canonical unit vectors.

Whenever $T$ is any $n$-dimensional simplex $T$,
we define the \textit{aspect shape measure} $\aspectratio(T)$,
and 
the \textit{algebraic shape measure} $\algebraicshapemeasure(T)$
by 
\begin{align}\label{math:shapemeasure}
    \aspectratio(T)
    := 
    \frac{ \diam(T) }{ \height(T) }
    ,
    \qquad 
    \algebraicshapemeasure(T)
    := 
    \sup_{ \varphi : \Delta^n \rightarrow T } 
    % \frac{ \sigma_{\max}( \Jacobian \varphi ) }{ \sigma_{\min}( \Jacobian \varphi ) }
    \matnorm{\Jacobian \varphi} \matnorm{ \Jacobian \varphi^{-1} }
    ,
\end{align}
where the last supremum is taken over all affine transformation from the reference $n$-simplex onto the $n$-simplex $T$. 
% DONE: Although this is fairly clear upon reading, we could specify in word that $\geometricshapemeasure(T)$ is (equivalent to) the usual shape regularity parameter, whereas $\volumeratio(\mathcal T)$ is the maximal ratio of volumes between neighboring elements.
When $\calT$ is an $n$-dimensional simplicial complex, we naturally define 
\begin{align}\label{math:shapemeasure:triangulation}
    \aspectratio(\calT) := \sup_{ T \in \subsimplex_{n}(\calT) } \aspectratio(T)
    ,
    \quad 
    \algebraicshapemeasure(\calT) := \sup_{ T \in \subsimplex_{n}(\calT) } \algebraicshapemeasure(T)
    .
\end{align}
We call these the aspect and algebraic shape measure, respectively, of the triangulation.

We gather a few relationships between geometric and algebraic entities and compare the different shape measures of a single simplex.

\begin{lemma}\label{lemma:measurerelationships}
    Let $T$ be an $n$-simplex and let $\varphi : \Delta^{n} \rightarrow T$ be an affine diffeomorphism from the reference $n$-simplex. Then 
    \begin{gather*} % First pair 
        \matnorm{ \Jacobian \varphi } 
        \leq 
        \Ceins{n} \cdot \diam(T),
        \qquad 
        \matnorm{ \Jacobian \varphi^{-1} } 
        \leq 
        \Czwei{n} \cdot \aspectratio(T) \height(T)^{-1}
        ,
        \\
        \frac{1}{\sqrt{2n}} \aspectratio(T) \leq \algebraicshapemeasure(T) \leq n \aspectratio(T).
    \end{gather*}
    Here, $\Ceins{n} = \Czwei{n} = \sqrt n$. 
\end{lemma}
\begin{proof}
    Let $\varphi : \Delta^{n} \rightarrow T$ be an affine transformation. 
	We abbreviate $M := \Jacobian\phi$ for its Jacobian. 
    We begin with observing that the largest $\ell^{2}$-norm of any column of $M$,
    which here denote by $c_{\max}(M)$, 
    equals the maximum of the quotient $\| M x \|_{\ell^{2}} / \| x \|_{\ell^{1}}$ over all non-zero $x \in \bbR^{n}$. 
    We also know that the diameter of $T$ is the length of its longest edge. 
    Our first pair of inequalities follows via standard comparisons of Euclidean norms:
    \begin{align*}
        \frac{\diam(T)}{\sqrt 2} \leq \matnorm{ M } \leq \sqrt{n} c_{\max}(M) \leq \sqrt{n} \diam(T).
    \end{align*}
    The columns of the matrix $M^{-1}$ are the gradients of the barycentric coordinates of the vertices of $T$,
    except for the vertex $\varphi(0)$. It immediately follows that 
    \begin{align*}
        \matnorm{ M^{-1} } \leq \sqrt{n} c_{\max}(M^{-1}) \leq \sqrt{n} \cdot \height(T)^{-1}
        .
    \end{align*}
    The smallest height in the reference simplex $\Delta^{n}$ is $\height_{\Delta} = 1/\sqrt{n}$, 
    whence $\height(T) \geq \matnorm{M^{-1}}^{-1} / \sqrt{n}$. This yields our second pair of inequalities. 
    Notice that $\height(T)^{-1} \leq \aspectratio(T) \diam(T)$.
    All relevant results follow.
\end{proof}

We are interested in the maximal ratio of volumes between face-neighboring $n$-simplices, written $\volumeratio(\calT)$,
and the ratio of the diameters of any intersecting simplices, written $\diameterratio(\calT)$. Formally, 
\begin{align}\label{math:volumeratio}
    \volumeratio(\calT) 
    := 
    \sup_{ \substack{ T, T' \in \subsimplex_{n}(\calT) \\ T \cap T' \in \subsimplex_{n-1}(\calT) } } 
    \frac{ \vol(T) }{ \vol(T') }
    ,
\end{align}
\begin{align}\label{math:diameterratio}
    \diameterratio(\calT) 
    := 
    \sup_{ \substack{ T, T' \in \subsimplex_{n}(\calT) \\ T \cap T' \neq \emptyset } }
    \frac{ \diam(T) }{ \diam(T') }
    .
\end{align}
Finally, whenever $T, T'$ are two $n$-simplices that share a common face $F$ of codimension $1$, we let $\Xi_{T,T'} : T \rightarrow T'$ denote the affine diffeomorphism 
that preserves $F$. We then define 
\begin{align}\label{math:reflectionmeasure}
    \reflectionestimate(\calT) 
    := 
    \sup_{ \substack{ T, T' \in \subsimplex_{n}(\calT) \\ T \cap T' \in \subsimplex_{n-1}(\calT) } }
    \matnorm{ \Jacobian \Xi_{T,T'} }
\end{align}
to be the maximum of the operator norm of the Jacobian of any such diffeomorphism. 
This indicator quantifies how much reflection across the shared face distorts the geometry.

\begin{lemma}\label{lemma:volumecomparison}
    Let $T_1$ and $T_2$ be two $n$-simplices that share a common face $F$. Then 
    \begin{align*}
        \diam(T_1)
        \leq 
        % \frac{ \diam(T_1)^{n} }{ n! \vol(T_1) }
        % \diam(F)
        % = 
        \aspectratio(T_1)
        \diam(F)
        ,
        \qquad 
        \frac{ \vol(T_1) }{ \vol(T_2) }
        \leq 
        % \frac{\diam(T_1)^{n}}{n! \cdot \vol(T_1)} 
        % \cdot 
        % \frac{\diam(T_2)^{n}}{n! \vol(T_2)} 
        % = 
        \aspectratio(T_1) \aspectratio(T_2)
        .
    \end{align*}
    If $\Xi : T_1 \rightarrow T_2$ is an affine diffeomorphism that is the identity over $F$, then 
    \begin{align*}
        \matnorm{ \Jacobian \Xi }
        \leq 
        ( 1 + \aspectratio(T_2) ) \aspectratio(T_1)
        % \Cvier{n} \geometricshapemeasure(T_2) \geometricshapemeasure(T_1) 
        .
    \end{align*}
%     Here, $\Cvier{n} \leq \Ceins{n} \Czwei{n}$. 
\end{lemma}

\begin{proof}
    The diameter of $F$ is at least as large as the height $h_S$ of some other vertex of $F$ in $T_{1}$. 
    Now, 
    \begin{align*}
%        \diam(T) \frac{ n! \vol(T) }{ \diam(T)^{n} } \leq h_S \leq \diam(F).
        \diam(T_{1}) \aspectratio(T_{1})^{-1} \leq \height(T_{1}) \leq h_S \leq \diam(F).
    \end{align*}
    The first estimate follows. 
    As for the second estimate,
    let $h_1$ and $h_2$ be the heights of $F$ in the simplices $T_1$ and $T_2$, respectively. 
    By the volume formula for simplices, $\vol(T_1) = h_1 \vol(F)/n$ and $\vol(T_2) = h_2 \vol(F)/n$, and thus $\vol(T_1)/\vol(T_2) = h_1/h_2$. Thus follows the second estimate: 
    \begin{align*}
        \frac{ \vol(T_1) }{ \vol(T_2) }
        = 
        \frac{ h_1 }{ h_2 }
        \leq
        \frac{ \diam(T_1) }{ h_2 }
        \leq
        \aspectratio(T_{1}) \frac{ \diam(F) }{ h_2 }
        \leq
        \aspectratio(T_{1}) \aspectratio(T_{2})
        .
    \end{align*}
    Lastly, let $\Xi : T_1 \rightarrow T_2$ be as stated. 
    Writing $h_0$ for the unit normal in direction $z_1$, we have 
    \begin{align*}
        \Xi(x) 
        = 
        x 
        - \frac{ \langle h_0, x \rangle }{ \langle h_0, z_1 \rangle } z_1
        + \frac{ \langle h_0, x \rangle }{ \langle h_0, z_1 \rangle } z_2
        ,
        \qquad 
        x \in \bbR^{n}
        .
    \end{align*}
    Its Jacobian is the rank-one perturbation of the linear projection along $z_1$ onto the hyperplane spanned by $F$. 
    By the Xu-Zikanatov estimate,
    \begin{align*}
        \matnorm{ \Jacobian \Xi }
        \leq 
        \frac{ \diam(T_1) }{ \height(T_1) }
        +
        \frac{ \diam(T_2) }{ \height(T_1) }
        \leq 
        \left( 1 + \aspectratio(T_2) \right)
        \aspectratio(T_1)
        .
    \end{align*}
    The desired estimates are shown.
\begin{comment}
    \mwl{Tighter estimate via geometry}
    Matrix inverse lemma:
    \begin{align*}
        A ( A + (z_2-z_1) e_1^t )^{-1}
        =
        \Id
        - 
        \left( 1 + e_1^{t} A^{-1} (z_2-z_1) \right)^{-1}
        (z_2-z_1) e_1^t \cdot A^{-1}
        =
        \Id
        - 
        \left( e_1^{t} A^{-1} z_2 \right)^{-1}
        (z_2-z_1) e_1^t \cdot A^{-1}
        .
    \end{align*}
    Putting in a multiple of $\alpha z_1$, we find 
    \begin{align*}
        \alpha z_1
        -
        \left( e_1^{t} A^{-1} z_2 \right)^{-1}
        (z_2-z_1) 
        .
    \end{align*}

@@

\begin{remark}
    While we will utilize $\diameterratio(\calT)$ at numerous places throughout the manuscript, 
	$\volumeratio(\calT)$ and $\reflectionestimate(\calT)$ will only be used throughout Section~\ref{section:gradient},
	the following section.  
	Lemma~\ref{lemma:volumecomparison} obviously shows that $\volumeratio(\calT)$ of~\eqref{math:volumeratio} is controlled by the shape measure. In a face-connected triangulation where we have an upper bound for the number of simplices sharing a vertex, 
    this Lemma also allows control of $\diameterratio(\calT)$ of~\eqref{math:diameterratio}.
\end{remark}

@@

\begin{corollary}
    Under the assumptions of Theorem~\ref{theorem:poincarefriedrichsestimate:grad},
    whenever $T_0, T_1, \dots, T_m$ is a face-path of $n$-simplices in $\calT$,
    we have 
    \begin{align*}
        \| w \|_{L^{p}(T_m)}
        &\leq 
        \sum_{\ell=1}^{m} 
        \volumeratio^{\frac{m-\ell}{p}}
        \min\left( 
        \left( 1 + \volumeratio(\calT)^{\frac q p} \right)^{\frac 1 q}
        C_{{\PF},\Minako_{F_\ell},p} 
        ,
        \left( 1 + \volumeratio(\calT)^{\frac q p} \reflectionestimate(\calT)^{q} \right)^{\frac 1 q}
        C_{\PF,T_{\ell},F_{\ell},p} 
        \right)
        \| \nabla u \|_{L^{p}(\Minako_{F_\ell})}
        \\&
        \qquad 
        \qquad 
        +
        \volumeratio^{\frac m p}
        C_{{\PF},T_{0},p}
        \| \nabla u \|_{L^{p}(T_{0})}
        .
    \end{align*}
\end{corollary}
\begin{proof}
    This follows by repeated application of the recursive estimate in Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.
\end{proof}
